---
title: "Untitled"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(lubridate)
library(tsibbledata)
library(janitor)
library(tidyverse)
library(ggplot2)
library(tsibble)
```

What is the pattern of bike hires over time (e.g. within a year, month, week, or day)?

#ok so what makes people rent a bike, or choose not to?
#off the top of my head, I can think of a couple: weather, being able to afford a taxi (after payday), being hungover.
#so it's probably not being able to afford it, since most of the rides are subscription--- although is that the right reasoning?
#do people cycle less in the winter? Is there a gender divide on that front? Women have a higher pain threshold so may stick around in winter.


#make a graph highlighting differences between men and women cyclists in the different seasons
#make a graph highlighting the differences between riders on weekdays and weekends
#make a graph highlighting the difference between riders after payday and before

Do bike hire patterns differ between bike rider demographics? (e.g. gender, type of trip, age)

Any other insights?


#reading in names and cleaning

```{r}
nyc_bikes_df <- nyc_bikes %>% 
  clean_names()

nyc_bikes_df
```

#giving extra columns for each of month, day, year.

```{r}
nyc_bikes_df <-
nyc_bikes_df %>% 
  mutate(year = year(start_time),
         month = month(start_time),
         week = week(start_time),
         day = day(start_time)
         ) %>%
  mutate(weekday = weekdays(start_time)) %>% 
  mutate(hour = hour(start_time),
         minute = minute(start_time),
         second = second(start_time)) %>% 
  mutate(time_start = format(start_time, format = "%H:%M:%S")) %>% 
  mutate(start_date = date(start_time))

nyc_bikes_df
```




```{r}
#summary of all hires on each individual day of the DF

nyc_bikes_df_summ <-
nyc_bikes_df %>% 
  index_by(start_date) %>% 
  summarise(hires = n()) 

nyc_bikes_df_summ

#summary of all hires on each month of the DF

#want to create a column that summarises how many women rented bikes per month, how many men hired bikes per month

nyc_bikes_df_summ_month <-
  nyc_bikes_df %>% 
  index_by(month) %>% 
  summarise(hires = n())
            #,
            #female = n(gender == "Female")) 
nyc_bikes_df_summ_month


#nyc_bikes_df_summ_month_gender <-
  nyc_bikes_df %>% 
  group_by(gender, type) %>% 
  summarise(count = n())
  
  

nyc_bikes_df_summ_month_gender


#summary of all hires on each day of the month in the DF (1st, 23rd etc)

nyc_bikes_df_summ_day <-
  nyc_bikes_df %>% 
  index_by(day) %>% 
  summarise(hires = n())

nyc_bikes_df_summ_day

#summary of all hires on each  of the DF. For use in bar or col graph

nyc_bikes_df_summ_day_before_10 <-
  nyc_bikes_df %>% 
  filter(hour <= 10) %>% 
  index_by(hour) %>% 
  summarise(hires = n())

nyc_bikes_df_summ_day_before_10
```


```{r}
nyc_bikes_df_summ %>%
  ggplot(aes(x = start_date, y = hires)) +
  geom_point() + 
  geom_line()

nyc_bikes_df_summ_month %>%
  ggplot(aes(x = month, y = hires)) +
  geom_point() + 
  geom_line() +
  scale_x_continuous(breaks = 1:12)

nyc_bikes_df_summ_day %>%
  ggplot(aes(x = day, y = hires)) +
  geom_point() + 
  geom_line() +
  scale_x_continuous(breaks = 1:31) +
  geom_smooth(method = lm)

```


 







































