---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(dplyr)
library(janitor)

```

```{r}
hb_geo_codes_14_19 <-
  read_csv("raw_data/health_board_geography_codes_and_labels_2014_01042019.csv") %>% 
  clean_names()

hb_geo_codes_14_19


health_board_incidences <-
  read_csv("raw_data/opendata_inc9418_hb.csv") %>% 
  clean_names()

health_board_incidences

region_incidences <-
  read_csv("raw_data/opendata_inc9418_region.csv") %>% 
  clean_names()

region_incidences

scotland_incidences <-
  read_csv("raw_data/opendata_inc9418_scotland.csv") %>% 
  clean_names() 

scotland_incidences

```


So hb_geo_codes_14_19 is simply the health board codes for each of the health boards in scotland. I have to do join it to health_board_incidences and then filter for NHS Borders
Or not even, just look up the hb code for NHS Borders in 1st and filter by that in the 2nd. 

```{r}
hb_geo_codes_14_19 %>% 
  filter(hb_name == "NHS Borders")
```


```{r}
health_board_incidences_Borders <-
health_board_incidences %>% 
  filter(hb == "S08000016") 

health_board_incidences_Borders

health_board_incidences_Borders %>% 
  distinct(easr_upper95pc_confidence_interval)

```

Questions I can think to ask, off the bat:

```{r}
health_board_incidences_Borders %>% 
  filter(sex == "All") %>% 
  ggplot(aes(x = year, y = incidences_all_ages)) +
  geom_line(color = "#800080") +
  facet_wrap(~ cancer_site)
```
So off the bat here I'm seeing that All Cancer Types may be making this data less accessible.

So I'm going to run the same code but filter out "All cancer types" to see what happens

```{r}
health_board_incidences_Borders %>% 
  filter(sex == "All",
         cancer_site != "All cancer types") %>% 
  ggplot(aes(x = year, y = incidences_all_ages)) +
  geom_line(color = "#800080") +
  facet_wrap(~ cancer_site)
```


```{r}
write_csv(health_board_incidences_Borders, "clean_data/health_board_incidences_Borders.csv")
```




```{r}
cancers_1 <-
unique(health_board_incidences_Borders$cancer_site)
```


```{r}
cancers_1
```

```{r}
health_board_incidences_Borders %>% 
  filter(sex == "All",
         cancer_site != "All cancer types") %>% 
  ggplot(aes(x = year, y = incidences_all_ages)) +
  geom_line(color = "#800080") + 
  facet_wrap(~ cancer_site)
```

```{r}
unique(health_board_incidences_Borders$sex)
```



```{r}
health_board_incidences_Borders %>% 
  filter(cancer_site == "Breast") %>% 
  group_by(year)
```


```{r}
health_board_incidences_Borders
```


```{r}
health_board_incidences_Borders %>% 
  tidyr::gather()
  filter(sex == "All",
         cancer_site != "All cancer types") %>% 
  ggplot(aes(x = year, y = incidences_all_ages)) +
  geom_line(color = "#800080") + 
  facet_wrap(~ cancer_site)
```

```{r}
unique(health_board_incidences_Borders$year)
```


```{r}
health_board_incidences_Borders %>% 
  filter(cancer_site != "All cancer types") %>% 
  group_by(cancer_site) %>% 
  arrange(desc(incidences_all_ages))
```


```{r}
health_board_incidences_Borders %>% 
  dplyr::select(-crude_rate_lower95pc_confidence_interval, -crude_rate_upper95pc_confidence_interval, -sex_qf, cancer_site_icd10code, -easr_lower95pc_confidence_interval, -easr_lower95pc_confidence_interval_qf, -easr_upper95pc_confidence_interval, -easr_upper95pc_confidence_interval_qf, -wasr_lower95pc_confidence_interval, -wasr_lower95pc_confidence_interval_qf, -wasr_upper95pc_confidence_interval, -wasr_upper95pc_confidence_interval_qf, -sir_lower95pc_confidence_interval, -sir_upper95pc_confidence_interval) %>% 
  pivot_longer(cols = c("incidences_all_ages", "crude_rate", "easr",),
               names_to = "metric",
               values_to = "value")
```

```{r}
read_csv("raw_data/borders_population.csv") %>% 
  clean_names() %>% 
  group_by(date_code) %>% 
  summarise(total = sum(value))
```


```{r}
borders_population <-
read_csv("raw_data/borders_population.csv") %>% 
  clean_names() %>% 
  filter(feature_code == "S08000016",
         age == "All",
         sex == "All") %>% 
  mutate(population = value) %>% 
  dplyr::select(-measurement, -units, -age, -sex, -value)

borders_population
```

```{r}
health_board_incidences_Borders_population <-
inner_join(borders_population, health_board_incidences_Borders, by = c("date_code" = "year")) %>% 
  dplyr::select(-crude_rate_lower95pc_confidence_interval, -crude_rate_upper95pc_confidence_interval, -sex_qf, cancer_site_icd10code, -easr_lower95pc_confidence_interval, -easr_lower95pc_confidence_interval_qf, -easr_upper95pc_confidence_interval, -easr_upper95pc_confidence_interval_qf, -wasr_lower95pc_confidence_interval, -wasr_lower95pc_confidence_interval_qf, -wasr_upper95pc_confidence_interval, -wasr_upper95pc_confidence_interval_qf, -sir_lower95pc_confidence_interval, -sir_upper95pc_confidence_interval) %>% 
  mutate(per_hundred_thousand = incidences_all_ages/population * 100000)

health_board_incidences_Borders_population
```

```{r}
health_board_incidences_Borders_population %>% 
  pivot_longer(cols = c("incidences_all_ages", "crude_rate", "easr",),
               names_to = "metric",
               values_to = "value") 
```


```{r}
health_board_incidences_Borders_population %>% 
  pivot_longer(cols = c("incidences_all_ages", "crude_rate", "easr",),
               names_to = "metric",
               values_to = "value") %>% 
  group_by(cancer_site) %>% 
  ggplot() +
  aes(x = date_code, y = value, fill = metric) +
  geom_line()
   
```

```{r}
health_board_incidences_Borders_lean <-
  health_board_incidences_Borders %>%  
  dplyr::select(-crude_rate_lower95pc_confidence_interval, -crude_rate_upper95pc_confidence_interval, -sex_qf, cancer_site_icd10code, -easr_lower95pc_confidence_interval, -easr_lower95pc_confidence_interval_qf, -easr_upper95pc_confidence_interval, -easr_upper95pc_confidence_interval_qf, -wasr_lower95pc_confidence_interval, -wasr_lower95pc_confidence_interval_qf, -wasr_upper95pc_confidence_interval, -wasr_upper95pc_confidence_interval_qf, -sir_lower95pc_confidence_interval, -sir_upper95pc_confidence_interval)
  
```


```{r}
health_board_incidences_Borders_lean %>% 
  filter(year >= 2005, 
         cancer_site != "All cancer types") %>% 
  group_by(cancer_site) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidences_all_ages), y = incidences_all_ages) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  geom_col(fill = "#800020") +
  scale_x_discrete() +
  labs(x = "Type of Cancer", y = "Incidences") +
  coord_flip()

```

So I'm going with the last ten years of data, which seems like more than enough for the topic.


```{r}
health_board_incidences_Borders_lean %>% 
  filter(year >= 2005, 
         cancer_site != "All cancer types",
         sex == "Male") %>% 
  group_by(cancer_site) %>% 
  arrange(desc(incidences_all_ages)) 
```


```{r}
health_board_incidences_Borders %>% 
  filter(cancer_site == "Non-melanoma skin cancer" | cancer_site ==  "Basal cell carcinoma of the skin" | cancer_site == "Squamous cell carcinoma of the skin" ,
         year >= 2005) %>% 
  summarise(sum = sum(incidences_all_ages))
```
```{r}
health_board_incidences_Borders %>% 
  filter(cancer_site == "Prostate",
         year >= 2005) %>% 
  summarise(sum = sum(incidences_all_ages))
```

```{r}
health_board_incidences_Borders %>% 
  filter(cancer_site == "Rectum and rectosigmoid junction" | cancer_site ==  "Colon" | cancer_site == "Colorectal cancer" ,
         year >= 2005) %>% 
  summarise(sum = sum(incidences_all_ages))
```







```{r}
sort(cancers)
```




```{r}
unique(health_board_incidences_Borders$cancer_site_icd10code)
```

```{r}
health_board_incidences %>% 
  filter(cancer_site_icd10code == "C50")
```
```{r}
health_board_incidences %>% 
  filter(grepl("C44", cancer_site_icd10code)) %>% 
  filter(cancer_site != "All cancer types") %>% 
  distinct(cancer_site_icd10code)

```
```{r}
health_board_incidences %>% 
  filter(grepl("C8", cancer_site_icd10code)) %>% 
  filter(cancer_site != "All cancer types") %>% 
  distinct(cancer_site_icd10code)

```


```{r}
health_board_incidences %>% 
  distinct(as.vector(cancer_site_icd10code)) %>% 
  arrange()
```



```{r}
health_board_incidences_Borders_icd <-
  health_board_incidences_Borders %>% 
  #filter out the more narrow ranges, there's colorectal 18-20 so i don't need 18 or 19-20. Same with c44, it includes both squamous and basal
  filter(cancer_site_icd10code != "C19-C20",
         cancer_site_icd10code != "C18",
         cancer_site_icd10code != "C92.1-C92.2",
         #skin cancer
         cancer_site_icd10code != "C44, M-8050-8078, M-8083-8084",
         cancer_site_icd10code != "C44, M-8090-8098",
         #leukemia
         cancer_site_icd10code != "C92.1-C92.2", 
         cancer_site_icd10code != "C01, C02.4, C05.1, C05.2, C09, C10",
         cancer_site_icd10code != "C00-C14",
         cancer_site_icd10code != "C01, C02.4, C05.1, C05.2, C09, C10",
         cancer_site_icd10code != "C01, C02.4, C05.1, C05.2, C09, C10",         
         cancer_site_icd10code != "C00-C14, C30-C32	",
         cancer_site_icd10code != "C00-C97, excluding C44	",
         #keeping 01-06, which is oral cancer
         cancer_site_icd10code != "C01-C02",
         cancer_site_icd10code != "C03-C06",
         #keeping 53-55, which is Uterine cancer
         cancer_site_icd10code != "C53",
         cancer_site_icd10code != "C54",
         #not sure what this is but removing duplicate not including the D numbers
         cancer_site_icd10code != "C70-C72, C75.1-C75.3",
         cancer_site_icd10code != "C91.0",
         cancer_site_icd10code != "C91.1",
         cancer_site_icd10code != "C92.0",


         
         
         )

health_board_incidences_Borders_icd %>% 
  distinct(cancer_site_icd10code) %>% 
  arrange(cancer_site_icd10code)
```

```{r}
health_board_incidences_Borders_icd %>% 
  filter(year >= 2005, 
         cancer_site != "All cancer types") %>% 
  ggplot() +
  aes(x = reorder(cancer_site, -incidences_all_ages), y = incidences_all_ages) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  geom_col(fill = "#800020") +
  scale_x_discrete() +
  labs(x = "Type of Cancer", y = "Incidences") +
  coord_flip()

```

```{r}
write_csv(health_board_incidences_Borders_icd, "clean_data/health_board_incidences_Borders_icd.csv")
```


```{r}

health_board_incidences_Borders_icd %>% 
  distinct(as.vector(cancer_site)) %>% 
  arrange()

cancer_sites_icd <-
sort(unique(health_board_incidences_Borders_icd$cancer_site))
```


```{r}
health_board_incidences_Borders_icd %>% 
  filter(year >= 2005,
         cancer_site != "All cancer types",
         sex == "All") %>% 
  group_by(cancer_site) %>% 
    summarise(incidences_2018 = sum(incidences_all_ages)) %>% 
  arrange(desc(incidences_2018)) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidences_2018), y = incidences_2018) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  geom_col(fill = "#800020") +
  scale_x_discrete() +
  labs(x = "Type of Cancer", y = "Since 2005") +
  coord_flip()
  
 
    
```

```{r}
health_board_incidences_Borders_icd %>% 
  filter(year == 2018,
         cancer_site != "All cancer types",
         sex == "All") %>% 
  group_by(cancer_site) %>% 
    summarise(incidences_2018 = sum(incidences_all_ages)) %>% 
  arrange(desc(incidences_2018)) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidences_2018), y = incidences_2018) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  geom_col(fill = "#800020") +
  scale_x_discrete() +
  labs(x = "Type of Cancer", y = "2018") +
  coord_flip()
  
 
    
```




```{r}
health_board_incidences_Borders_icd %>% 
  filter(year == 2018,
         sex == "All") %>% 
  group_by(cancer_site) %>% 
    summarise(incidences_since_2005 = sum(incidences_all_ages)) %>% 
  arrange(desc(incidences_since_2005))
   
```

```{r}
health_board_incidences_Borders_icd %>% 
  filter(cancer_site == "Prostate")
```

```{r}
health_board_incidences_Borders_icd %>% 
  filter(year == 2018,
         cancer_site != "All cancer types",
         sex == "Female") %>% 
  group_by(cancer_site) %>% 
    summarise(incidences_2018 = sum(incidences_all_ages)) %>% 
  arrange(desc(incidences_2018)) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidences_2018), y = incidences_2018) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  geom_col(fill = "#800020") +
  scale_x_discrete() +
  labs(x = "Type of Cancer", y = "2018") +
  coord_flip()
  
 
```

```{r}
health_board_incidences_Borders_icd %>% 
  filter(year == 2018,
         cancer_site != "All cancer types",
         sex == "Male") %>% 
  group_by(cancer_site) %>% 
    summarise(incidences_2018 = sum(incidences_all_ages)) %>% 
  arrange(desc(incidences_2018)) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidences_2018), y = incidences_2018) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  geom_col(fill = "#800020") +
  scale_x_discrete() +
  labs(x = "Type of Cancer", y = "2018") +
  coord_flip()
  
 
```



























