---
title: "R Notebook"
output: html_notebook
---

1 MVP

1.1 Hypothesis testing - practical

#You work for an animal conservation organisation and are looking to do some investigating into some of the animals to provide evidence for some projects you are looking to run.

#In this homework we’ll use the msleep data set provided in the ggplot package. It concerns the sleeping patterns of various types of mammal.


#Explore the dataset and familiarise yourself with it.


```{r}
library(tidyverse)
library(dplyr)
library(janitor)
library(infer)
library(fastGraph)
```


```{r}
msleep <- msleep %>% 
  clean_names()

msleep

msleep <-
msleep %>% 
  add_row(name = "Jabberwocky", sleep_total = 7)
```


```{r}

```




#Jabberwockies sleep for around 7 hours a night, on average. Perform an appropriate statistical test to determine whether the mean sleep_total in the sampled population of animal types differs from the typical value for jabberwockies.

Do animals have a mean(sleep_total) significantly different from 3.96?

α level: 0.05

H0: the mean of sleep_total in the msleep dataframe does NOT differ from the average 7 hours that Jabberwockies spend sleeping   (μsleep_total_mean = 7)
Ha: the mean of sleep_total in the msleep dataframe DOES differ from the average 7 hours that Jabberwockies spend sleeping (μsleep_total_mean ≠ 7)

```{r}
#simulating a sample of the animals, using replace to generate more samples than there are in the df
sample_200 <- msleep %>% 
  rep_sample_n(size = 200, reps = 1, replace = TRUE)
```

```{r}
sleep_total_mean <- sample_200 %>% 
  summarise(mean_sleep_total = mean(sleep_total))

sleep_total_mean
```

```{r}
sleep_total_mean_jabberwocky <- sample_200 %>% 
  filter(name == "Jabberwocky") %>% 
  summarise(mean_sleep_total = mean(sleep_total)) 

sleep_total_mean_jabberwocky
```

```{r}
std_errs <- sample_200 %>% 
  summarise(se_sleep_total_mean = sd(sleep_total_mean))

std_errs
```

```{r}  
null_distribution <- sample_200 %>% 
  specify(response = sleep_total) %>% 
  hypothesize(null = "point", mu = 7) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean")

null_distribution %>% 
  visualise(bins = 30) +
  shade_p_value(obs_stat = sleep_total_mean$mean_sleep_total, direction = "less")


p_value_1 <- null_distribution %>% 
  get_p_value(sleep_total_mean_jabberwocky, direction = "less")
p_value_1


#p value = 0, therefore we should reject the null hypothesis
```




#Perform an appropriate statistical test to determine whether omnivores sleep for significantly longer than herbivores, on average.

```{r}
sample_200
```



```{r}
null_distribution_2 <- sample_200 %>% 
  specify(sleep_total ~ vore) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi"))

null_distribution_2

#I keep getting this error and I can't for the life of me figure out why it's doing that- haven't I already separated the variable into two levels by creating a vector?

#Error: Statistic is based on a difference or ratio; the explanatory variableshould have two levels.

```





#Perform an appropriate statistical test to determine whether the proportion of domesticated animal types in the population of animal types is greater than 5%.
#[Hint - think about creating an is_domesticated variable for the analysis]

```{r}
msleep <-
msleep %>% 
  mutate(is_domesticated = ifelse(conservation == "domesticated", "Yes", "No"))

msleep
```

```{r}
null_distribution_3 <- msleep %>% 
  specify(name ~ is_domesticated) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 5000, type = "permute") %>% 
  calculate(stat = "diff in prop", order = c("Yes", "No"))

null_distribution_3
```





1.2 Hypothesis Testing 

1. Coffee shop

H0:μawareness_of_shop > 0.4
Ha:μawareness_of_shop < 0.4

H0: over 40% percent of the people in the town know the coffee shop
Ha: under 40% percent of the people in the town know the coffee shop

I would use a permutation to generate the null distribution(?)



2. Website A/B test




H0: website users overall are more likely to click on the banner if positioned at the top of the page
Ha: website users overall are less likely to click on the banner if positioned at the top of the page




1.2.2




























