---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
output: html_notebook
---


#1 MVP
#This is a mini homework for people doing the PDA. You are going to continue working on your week 2 weekend homework.

#We want you to use the skills you have learnt this week to visualise your meteorite data. We do not need anything fancy - just a simple plot that illustrates part of your cleaned data.





```{r}
library(tidyverse)
library(ggplot2)
```

1.2.1

```{r}
meteorite_landings_clean <-
  read_csv("data/meteorite_landings_clean.csv")
```


```{r}
ggplot(meteorite_landings_clean) +
  aes(x = year, y = mass_in_grams) +
  geom_line() +
  labs(x = "\n Year",
       y = "\n Mass In Grams") 
```
```{r}
ggplot(meteorite_landings_clean_over_10_mil) +
  aes(x = year, y = mass_in_grams, colour = fall_status) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Meteor weight by year, split by fall status") +
  labs(mass_in_grams = "Mass in Grams")
```

```{r}
ggplot(meteorite_landings_clean_under_10_mil_last_twenty_years) +
  aes(x = year, y = mass_in_grams, colour = fall_status) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Meteor weight by year, split by fall status") +
  labs(mass_in_grams = "Mass in Grams")
```




```{r}
meteorite_landings_clean_under_10_mil <-
  meteorite_landings_clean %>% 
  filter(mass_in_grams < 10000000)
```

```{r}
meteorite_landings_clean_over_10_mil <-
  meteorite_landings_clean %>% 
  filter(mass_in_grams > 10000000)
```


```{r}
meteorite_landings_clean_under_10_mil_last_hundred_years <-
  meteorite_landings_clean %>% 
  filter(mass_in_grams < 10000000) %>% 
  filter(year > 1921)
```


```{r}
meteorite_landings_clean_under_10_mil_last_twenty_years <-
  meteorite_landings_clean %>% 
  filter(mass_in_grams < 10000000) %>% 
  filter(year > 2001)
```






```{r}
library(raster)
library(rgdal)
library(ggplot2)
library(broom)
library(RColorBrewer)
library(rgeos)
library(dplyr)
library(tidyverse)
```



figure out how to divide the four corners by latitude and longitude
add up the sum of each of the areas
place on a bar chart



```{r}
meteorite_landings_clean %>% 
  mutate(hemisphere = if_else(latitude > 0, "North", "South")) %>% 
  head(mass_in_grams, 100)
```
```{r}
meteorite_landings_clean_top_100 <-
meteorite_landings_clean_hemisphere %>% 
  arrange(desc(mass_in_grams)) %>% 
  head(100)
```



```{r}
ggplot(meteorite_landings_clean_hemisphere) +
  aes(x = year, y = mass_in_grams, colour = hemisphere) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Meteor weight by year, split by hemisphere") +
  labs(mass_in_grams = "Mass in Grams")
```



```{r}
ggplot(meteorite_landings_clean_hemisphere_over_1_tonne) +
  aes(x = year, y = mass_in_grams, colour = hemisphere) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Meteor weight by year, split by hemisphere") +
  labs(mass_in_grams = "Mass in Grams")
```



```{r}
meteorite_landings_clean_hemisphere <-
meteorite_landings_clean %>% 
  mutate(hemisphere = if_else(latitude > 0, "North", "South"))
```


```{r}
meteorite_landings_clean_top_100 <-
meteorite_landings_clean_hemisphere %>% 
  arrange(desc(mass_in_grams)) %>% 
  head(100)
```


```{r}
ggplot(meteorite_landings_clean_top_100) +
  aes(x = year, y = mass_in_grams, colour = hemisphere) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("100 Largest Meteors By Year, Split By Hemisphere") +
  labs(y = "\n Mass in Grams",
       x = "\n Year",
       colour = "Hemisphere") +
  scale_colour_manual(breaks = c("North", "South"),
                      values = c("darkblue", "red"))
```
































































































































































































































































































































































































































































































































